---
title: "Home Work 2"
author: "Nicholas Nasser (UT EID - nan823)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(kableExtra)
profs = read.csv("profs.csv")
bikeshare = read.csv("bikeshare.csv")
```

## Problem 1

#### Part A

```{r, fig.align = "center", fig.height = 4, fig.width = 6}

ggplot(profs) + geom_histogram(aes(x = eval), color = "black", fill = "palegreen3", binwidth = 0.1) + labs(title = "Course Evaluation Scores for UT Professors", x = "Average Teaching Evaluation Score (Scale of 1-5)", y = "Count")

```

&nbsp;

The left-skewed histogram shown above has a mean score of `r round(mean(profs$eval), digits = 2)` and a standard deviation of `r round(sd(profs$eval), digits = 2)`. Based on these summary statistics, we can see that professors at the university are providing an overall satisfying course experience for students. That being said, the presence of scores near 2-2.5/5 indicate that some students are unhappy with a handful of professors.

\newpage

#### Part B

```{r, fig.align = "center", fig.height = 4, fig.width = 6}

ggplot(profs) + geom_boxplot(aes(x = native, y = eval), fill = "khaki3") + labs( title = "Course Evaluation Scores for UT Professors by Native Speaker Status", x = "Native Speaker of English", y = "Average Score (Scale of 1-5)")

profs_eng = filter(profs, native == "yes")
profs_non_eng = filter(profs, native == "no")
```

&nbsp;

Native English Speaking Professors:

```{r}

fivenum(profs_eng$eval)

```

Non-Native English Speaking Professors:

```{r}

fivenum(profs_non_eng$eval)

```

&nbsp;

After observing the side by side box plots shown above, it is clear that the average evaluation scores for native English speaking professors were higher than they were for non-native English speaking professors in terms of first quartile, median, third quartile, and maximum values, with the exception being the minimum values. This indicates that overall, students feel they benefit more from taking classes with native English speaking professors.

\newpage

#### Part C

```{r, fig.align = "center", fig.height = 6, fig.width = 6}

ggplot(profs) + geom_histogram(aes(x = eval), color = "black", fill = "pink2", binwidth = 0.1) + labs(title = "Course Evaluation Scores for UT Professors by Gender", x = "Average Teaching Evaluation Score (Scale of 1-5)", y = "Count") + facet_wrap(~gender, nrow = 2)

by_gender = group_by(profs, gender)
mean_sd_bygender = summarize(by_gender, mean_eval_score = round(mean(eval), digits = 2), standard_deviation = round(sd(eval), digits = 2))

mean_sd_bygender %>%
  kbl(caption = "Mean and Standard Deviation of Course Evaluation Scores by Gender") %>%
  kable_styling(latex_options = "hold_position")

```
&nbsp;

When looking at course evaluation scores for professors by gender, we see that male professors have a slightly higher average score than their female counterparts - 4.07/5 to 3.90/5 respectively. The standard deviation from the mean was 0.56 for male professors, and 0.54 for female professors. Average ratings of about 4.0/5 with a standard deviation of about 0.55 for both male and female professors demonstrate that there is not much correlation between a professor's gender and his or her evaluation score; students do not think much about - if it all - gender when giving their professors evaluation scores.

\newpage

#### Part D

```{r, fig.align = "center", fig.height = 4, fig.width = 6}

ggplot(profs) + geom_point(aes(x = beauty, y = eval)) + labs(title = "Professor Physical Attractiveness vs. Course Evaluation Scores", x = "Physcial Attractiveness (Deviation from Average)", y = "Average Score (Scale of 1-5)")

```

&nbsp;

Calculating the correlation between physical attractiveness of professors and the course evaluation scores they received reveals a low positive correlation coefficient of `r round(cor(profs$beauty, profs$eval), digits = 2)`. This low coefficient value as well as the random location of points on the scatter plot display a lack of any significant association between the two variables, meaning it is unlikely that students are rating professors based on their physical attractiveness.

\newpage

## Problem 2

#### Plot A

```{r, fig.align = "center", fig.height = 4, fig.width = 6}

average_hourly = group_by(bikeshare, hr) %>%
  
  summarise(average_rentals = mean(total)) %>%
  
  round(2)

ggplot(average_hourly) + geom_line(aes(x = hr, y = average_rentals)) + labs(title = "Average Hourly Bike Rentals for All Hours of the Day", x = "Hour of the Day (0-23)", y = "Average Hourly Bike Rentals (Total)")

```
&nbsp;

The line graph above plots the averages of total bike rentals for each hour of the day across two years (2011 and 2012) in Washington D.C. The x-axis represents the hours of the day, with 0 being 12:00 AM and 23 being 11:00 PM. The y-axis represents the average of total bikes rented during any given hour. Looking at the plot, we see that the peak times for bike rentals occur at around 7:30 AM and 5:30 PM, indicating increased demand for bike rentals during rush hours. We also see a dip in demand for bike rentals early in the mornings as well as late at night, when most people are usually at home or sleeping. Based on these observations, we can conclude that demand for rental bikes is partially reliant on when people are going to work.

\newpage

#### Plot B

```{r, fig.align = "center", message = FALSE}

average_hourly_f = group_by(bikeshare, hr, workingday) %>%
  
  summarise(average_rentals = mean(total)) %>%
  
  round(2)

workingday_facet = c(`0` = "Not a Working Day", `1` = "Working Day")

ggplot(average_hourly_f) + geom_line(aes(x = hr, y = average_rentals)) + facet_wrap(~workingday, labeller = as_labeller(workingday_facet)) + labs(title = "Average Hourly Bike Rentals for All Hours of the Day by Working Day", x = "Hour of the Day (0-23)", y = "Average Hourly Bike Rentals (Total)")

```
&nbsp;

The faceted line graphs above represent the same data and have the same axes as the line graph shown previously, with the only difference being that now the data is separated by whether or not it was a working day. Working days include all days that were not holidays or weekends. Observing these plots reveals that demand for rental bikes is in fact partially reliant on when people are commuting to and from work, as was determined from the previous plot. However, we now also know that a large portion of the bikes that were rented in between rush hours can be attributed to people going out leisurely on weekends and/or holidays during the middle of the day. This is why we see a much larger nosedive in demand for bikes in between rush hours during working days than we do on the original plot.

\newpage

#### Plot C

```{r, fig.align = "center", message = FALSE}

ridership_9am = filter(bikeshare, hr == "9") %>%
  
  group_by(weathersit, workingday) %>%
  
  summarise(average_rentals = mean(total)) %>%
  
  round(2)

ggplot(ridership_9am) + geom_col(aes(x = weathersit, y = average_rentals), color = "black", fill = "tomato2") + facet_wrap(~workingday, labeller = as_labeller(workingday_facet)) + labs(title = "Weather Situation vs. Ridership at 9 AM by Working Day ", x = "Weather Situation (Severity)", y = "Average Ridership")

```
&nbsp;

The bar plots above display the relationship between weather conditions and average ridership only during the 9 AM hour, and are faceted based on whether or not it was a working day. For a bit of context, here are descriptions for each of the four possible weather situations:

&nbsp;

1 - Clear, Few clouds, Partly cloudy, Partly cloudy

&nbsp;

2 -  Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist

&nbsp;

3 - Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds

&nbsp;

4 -  Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog

&nbsp;

As you can see, the fourth weather situation is not present on the bar plot because there were no occurrences of it during the 9 AM hour on any day in the data set. Looking at each bar plot, we can see that overall ridership levels at 9 AM are higher during working days. However, the more important fact to look at is that as the weather becomes more severe, less people use the bike share system regardless of whether it is a working day or not. We can simply conclude from this observation that weather/environmental factors play a large role in determining demand for rental bikes. 



